\huge \textbf{Unshuffling Permutations} \vspace{1.5em}

### 2016-04-08 slides/slides.tex
Simplifications

\uncover<2->{%
To take into account multiplicities, we consider $\shuffle$ as a linear
product
\begin{equation*}
    \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
\end{equation*}
on $\QQ[A^*]$, the $\QQ$-linear span of words defined by
\begin{align*}
    u \shuffle \epsilon & := u =: \epsilon \shuffle u, \\
    (ua \shuffle vb) & := (ua \shuffle v)b + (u \shuffle vb)a.
\end{align*}}
%
\uncover<3->{%
\begin{block}{Example}
    \vspace{-1em}
    \begin{equation*}\begin{split}
        \textcolor{RoyalBlue}{\La\Lb} \shuffle \textcolor{BrickRed}{\Lb\La}
        & =
        \textcolor{RoyalBlue}{\La\Lb}\textcolor{BrickRed}{\Lb\La}
        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb}%
        \textcolor{RoyalBlue}{\Lb}\textcolor{BrickRed}{\La}
        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb\La}%
        \textcolor{RoyalBlue}{\Lb}
        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La\Lb}%
        \textcolor{BrickRed}{\La}
        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La}%
        \textcolor{BrickRed}{\La}\textcolor{RoyalBlue}{\Lb}
        + \textcolor{BrickRed}{\Lb\La}\textcolor{RoyalBlue}{\La\Lb} \\
        &
        \uncover<4->{=
        2 \La\Lb\Lb\La + \La\Lb\La\Lb + \Lb\La\Lb\La + 2 \Lb\La\La\Lb}
    \end{split}\end{equation*}
    \vspace{-1.5em}
\end{block}}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Combinatorial products and structure coefficients}
Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
$\QQ$-linear span of $C$.
\medskip

A binary linear product
\begin{equation*}
    \cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C].
\end{equation*}
on $\QQ[C]$ describes a way to \alert{combine} two objects of
$C$ to obtain \alert{bigger} objects.
\medskip

In a general way, for any $x, y \in C$,
\begin{equation*}
    x \cdot y = \sum_{z \in C} \lambda_{x, y}^z \; z.
\end{equation*}
\medskip

A product is therefore wholly encoded by the coefficients
$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
\alert{structure coefficients}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Combinatorial coproducts}
In a dual way, a linear coproduct
\begin{equation*}
    \Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C].
\end{equation*}
on $\QQ[C]$ describe a way to \alert{break} one object of $C$ into
\alert{two smaller} pieces.
\medskip

In a general way, for any $z \in C$,
\begin{equation*}
    \Delta(z) = \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

A coproduct is therefore wholly encoded by the coefficients
$\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$, called
\alert{structure coefficients}.
\end{frame}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Equivalence between products and coproducts}
\begin{block}{Key idea}
    Knowing how to break a combinatorial object is equivalent
    to knowing how to combine these.
\end{block}
\medskip

Indeed, if $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$ is a product with
structure coefficients $\lambda_{x, y}^z \in \QQ$, $x, y, z \in C$,
we can define its \alert{dual coproduct} $\Delta$ by setting
\begin{equation*}
    \Delta(z) := \sum_{x, y \in C}  \lambda_{x, y}^z \;  x \otimes y.
\end{equation*}
\medskip

Conversely, if $\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$ is
a coproduct with structure coefficients $\lambda_{x, y}^z \in \QQ$,
$x, y, z \in C$, we can define its \alert{dual product} $\cdot$ by
setting
\begin{equation*}
    x \cdot y := \sum_{z \in C} \lambda_{x, y}^z \; z.
\end{equation*}
\end{frame}

>>>>

\uncover<2->{%
To take into account multiplicities, we consider $\shuffle$ as a linear
product
\begin{equation*}
    \shuffle : \QQ[A^*] \otimes \QQ[A^*] \to \QQ[A^*]
\end{equation*}
on $\QQ[A^*]$, the $\QQ$-linear span of words defined by
\begin{align*}
    u \shuffle \epsilon & := u =: \epsilon \shuffle u, \\
    (ua \shuffle vb) & := (ua \shuffle v)b + (u \shuffle vb)a.
\end{align*}}
%
\uncover<3->{%
\begin{block}{Example}
    \vspace{-1em}
    \begin{equation*}\begin{split}
        \textcolor{RoyalBlue}{\La\Lb} \shuffle \textcolor{BrickRed}{\Lb\La}
        & =
        \textcolor{RoyalBlue}{\La\Lb}\textcolor{BrickRed}{\Lb\La}
        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb}%
        \textcolor{RoyalBlue}{\Lb}\textcolor{BrickRed}{\La}
        + \textcolor{RoyalBlue}{\La}\textcolor{BrickRed}{\Lb\La}%
        \textcolor{RoyalBlue}{\Lb}
        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La\Lb}%
        \textcolor{BrickRed}{\La}
        + \textcolor{BrickRed}{\Lb}\textcolor{RoyalBlue}{\La}%
        \textcolor{BrickRed}{\La}\textcolor{RoyalBlue}{\Lb}
        + \textcolor{BrickRed}{\Lb\La}\textcolor{RoyalBlue}{\La\Lb} \\
        &
        \uncover<4->{=
        2 \La\Lb\Lb\La + \La\Lb\La\Lb + \Lb\La\Lb\La + 2 \Lb\La\La\Lb}
    \end{split}\end{equation*}
    \vspace{-1.5em}
\end{block}}
\end{frame}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{frame} \frametitle{Combinatorial products and coproducts}
Let $C$ be a set of combinatorial objects and $\QQ[C]$ be the
$\QQ$-linear span of $C$.

\uncover<2->{%
\begin{block}{Key idea}
    Knowing how to \alert{break} combinatorial objects explains
    how to \alert{combine} these.
\end{block}
\medskip}

\uncover<3->{%
To define a product $\cdot : \QQ[C] \otimes \QQ[C] \to \QQ[C]$,
it is some cases more convenient to start by defining a \alert{coproduct}
$\Delta : \QQ[C] \to \QQ[C] \otimes \QQ[C]$.
\medskip}

\uncover<4->{%
Every coproduct fits into the general form
\begin{equation*}
    \Delta(z)
    = \sum_{x, y \in C} \textcolor{RoyalBlue}{\lambda_{x, y}^z} \;
    (x \otimes y)
\end{equation*}
where the $\textcolor{RoyalBlue}{\lambda_{x, y}^z}$ are structure
coefficients of $\QQ$.
\medskip}

\uncover<5->{%
Then, $\Delta$ leads by duality to the definition of $\cdot$ by
\begin{equation*}
    x \cdot y
    := \sum_{z \in C} \textcolor{RoyalBlue}{\lambda_{x, y}^z} \; z.
\end{equation*}}
\end{frame}
